---
title: "Practical Machine Learning"
output: html_document
---
#Intro
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

##Libraries utilized
```{r,results='hide'}
library(caret); library(rpart)
library(rattle); library(rpart.plot)
library(randomForest); library(RColorBrewer)
library(corrplot)
```

##Load the training and test data sets
```{r}
training_data <- read.csv("./pml-training.csv")
testing_data <- read.csv("./pml-testing.csv")

dim(training_data);dim(testing_data)
```

##Clean up training data set 
This is done by removing missing values
```{r}
clean_train <- training_data
# pull out the "classe" field
train_classe <- as.data.frame(clean_train[, max(length(clean_train))])
colnames(train_classe) <- "classe"

# function to remove unnecessary columns
remove <- function (x) {
  remove <- grepl("^X|timestamp|window", names(x))
}

clean_train <- clean_train[,colSums(is.na(clean_train)) == FALSE]
# applying the remove function  to the new clean data set
clean_train <- clean_train[,!remove(clean_train)]
# converting columns to numeric values
clean_train <- clean_train[, sapply(clean_train, is.numeric)]
# add the "classe" values back in
clean_train <- cbind(clean_train, train_classe)
```
##Clean up testing data set by removing missing values
This is done by filtering down to complete objects
```{r}
clean_test <- testing_data
# pulls out the grades or "classe"
clean_test <- clean_test[, -max(length(clean_test))]
clean_test <- clean_test[, colSums(is.na(clean_test)) == 0]
# function to remove unnecessary columns

remove <- function (x) {
  remove <- grepl("^X|timestamp|window", names(x))
}
# applying the remove function
clean_test <- clean_test[,!remove(clean_test)]
#Convert columns to numeric values
clean_test <- clean_test[, sapply(clean_test, is.numeric)]

```

##Set the seed 
This is so that the results can be reproduced
```{r}
set.seed(532)
```

##Partition training and cross validation sets
```{r}
inTrain <- createDataPartition(y = clean_train$classe, p = .60, list = FALSE)
```

The data was split as follows. The initial 60% given to training data remaining 40% went to cross_validate set,
```{r}
clean_train <- clean_train[inTrain,]
cross_validate <- clean_train[-inTrain,]
#clean_train

plot(clean_train$classe, col="blue", main="Levels of Classe", xlab="classe levels", ylab="Frequency")
```
#Build the model
RandomForest was chosen due to its ability to produce fairly strong predictions.
```{r}

model <- train(classe ~ ., data = clean_train, method = "rf", 
               trControl = trainControl(method = "cv", 5), 
               ntree = 250)
model
```

##Predictions based on cross validated data
The model is now applied to the cross_validate data set.
```{r}
predict_it <- predict(model,cross_validate)

confusionMatrix(cross_validate$classe,predict_it)
```

Check the accuracy of the model
```{r}
model.accuracy <- postResample(predict_it, cross_validate$classe)
model.accuracy
```

Out of sample error
```{r}
samp.error <- 1 - 
  as.numeric(confusionMatrix(
    cross_validate$classe, 
    predict_it)$overall[1])
samp.error


clean_test$prediction <- predict(model, clean_test)
clean_test$prediction
```

## Decision Tree
```{r}
tree <- rpart(classe ~ ., data = clean_train, method = "class") 
fancytree <- fancyRpartPlot(tree)
basictree <- prp(tree)
```

## Prediction output
```{r}
#pml_write_files = function(x){
#  n = length(x)
#  for(i in 1:n){
#    filename = paste0("problem_id_",i,".txt")
#    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
#  }
#}
print(clean_test$prediction)

```
## Knitting function

```{r}
#library(knitr)

#build.report <- function(x) {
#  knit2html(x, "project.html")
#}
```
